---
title: "Rodenkirchen Straßenunfälle Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    fig_mobile: TRUE
    theme: 
      version: 4
      bootswatch: flatly
      base_font: 
        google: Prompt
      heading_font:
        google: Sen
---


```{r setup, include=FALSE}

library(flexdashboard)
library(leaflet)
library(sf)
library(dplyr)
library(crosstalk)

# load the data from the directory:

load("roki_dashboard_data.RData")

# 1. Transform spatial data from local CRS (EPSG:25832) to WGS84 (EPSG:4326) for Leaflet
# Ensure roki_clean is loaded in your environment (e.g., loaded from a GeoPackage or RDS file)
roki_clean1 <- st_transform(roki_clean, 4326)

# 2. Wrap the sf object in SharedData for client-side filtering without a Shiny server
sd <- SharedData$new(roki_clean1)
```



<style>
/* 1. Sophisticated Slate Sidebar & Header Styling */
body, .section.sidebar {
  background-color: #2c3e50 !important; /* Deep modern slate */
  color: #ecf0f1 !important;
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Roboto, "Helvetica Neue", Arial, sans-serif;
}

.navbar.navbar-inverse {
  background-color: #1a252f !important; /* Darker charcoal header */
  border-color: #1a252f !important;
}

/* 2. Refined Sidebar Spacing & Clean Padding */
.section.sidebar {
  padding: 20px 18px !important;
  border-right: 1px solid rgba(255, 255, 255, 0.1);
}

/* 3. Sleek Form Controls & Inputs */
.selectize-input, .form-control {
  background-color: #ffffff !important;
  border: 1px solid #cbd5e1 !important;
  border-radius: 6px !important;
  box-shadow: none !important;
  color: #1e293b !important;
}

/* Dropdown Arrow Customization */
.selectize-input::after {
  content: "▼" !important;
  font-size: 9px;
  position: absolute;
  right: 14px;
  top: 50%;
  transform: translateY(-50%);
  color: #64748b;
  pointer-events: none;
}
.selectize-input {
  padding-right: 30px !important;
  position: relative;
}

/* 4. Polished Sidebar Section Headings */
.section.sidebar .chart-title, .section.sidebar h3 {
  font-size: 13px !important;
  font-weight: 600 !important;
  letter-spacing: 0.5px;
  text-transform: uppercase;
  color: #94a3b8 !important;
  margin-top: 15px;
  margin-bottom: 6px;
  border-bottom: none !important;
}

/* 5. Soften Map Container Borders */
.leaflet-container {
  border-radius: 4px;
}

/* 6. Fix ionRangeSlider Track Inversion */
.irs-line {
  background: #1a252f !important;
  border: none !important;
}

.irs-bar {
  background: #ffffff !important;
  border: none !important;
}

.irs-from, .irs-to, .irs-single {
  background: #1a252f !important;
  color: #ecf0f1 !important;
}
</style>


Sidebar {.sidebar data-width=220}
-----------------------------------------------------------------------



```{r echo=FALSE}
# Control 1: Year Filter (Scalable for future year-over-year updates)
filter_checkbox(
  id = "year_filter", 
  label = "Select Year(s):", 
  sharedData = sd, 
  group = ~as.factor(year)
)

# Control 2: Month Filter (Using ordered factor 'Months')
filter_select(
  id = "month_filter", 
  label = "Select Month:", 
  sharedData = sd, 
  group = ~Months
)

# Control 3: Time-of-Day Filter (Granular hour slider from 0 to 23)
filter_slider(
  id = "hour_filter", 
  label = "Hour of Day:", 
  sharedData = sd, 
  column = ~hour, 
  step = 1, 
  sep = ""
)

# Control 4: Lighting Condition Filter
filter_select(
  id = "light_filter", 
  label = "Lighting Condition:", 
  sharedData = sd, 
  group = ~light_cond
)

# Control 5: Road Surface Condition Filter
filter_select(
  id = "road_filter", 
  label = "Road Surface Condition:", 
  sharedData = sd, 
  group = ~road_cond
)
```

---




Created by **Armond Angeles** (*Raster Ninja*). 

<div style="font-size: 13px; color: #94a3b8; margin-top: 15px; border-top: 1px solid rgba(255,255,255,0.1); padding-top: 10px;">
  <strong>Data Source:</strong> <a href="https://www.opengeodata.nrw.de/produkte/" target="_blank" style="color: #38bdf8; text-decoration: none;">OpenGeodata NRW</a>
</div>


```{r}
# load the data from the directory:

load("roki_dashboard_data.RData")

# Example summary variables based on your dataset columns
# (Adjust column names like `sektion` or `severity` to match your actual dataset)
total_fatalities <- sum(roki_clean$severity == "Fatal", na.rm = TRUE)
total_severe <- sum(roki_clean$severity == "Severe Injury", na.rm = TRUE)
total_slight <- sum(roki_clean$severity == "Slight Injury", na.rm = TRUE)
```

Row {data-height=5}
-----------------------------------------------------------------------

### Fatalities {.value-box}
```{r}


valueBox(
  value = total_fatalities,
  caption = "Getötete (Todesfälle)",
  icon = "fa-exclamation-triangle",
  color = "danger"
)
```

### Severe Injuries
```{r}
valueBox(
  value = total_severe,
  caption = "Schwerverletzte",
  icon = "fa-user-injured",
  color = "warning"
)
```

### Slight Injuries
```{r}
valueBox(
  value = total_slight,
  caption = "Leichtverletzte",
  icon = "fa-user",
  color = "info"
)
```



Row {data-height=80}
-------------------------------------

### Accident Map
```{r echo=FALSE}

leaflet(sd) %>%
  # 1. Google Road (Roadmap) Basemap
  addTiles(
    urlTemplate = "https://{s}.google.com/vt/lyrs=m&x={x}&y={y}&z={z}",
    options = tileOptions(subdomains = c('mt0', 'mt1', 'mt2', 'mt3')),
    attribution = '&copy; <a href="https://maps.google.com">Google Maps</a>',
    group = "Google Road"
  ) %>%
  # 2. Google Satellite Basemap
  addTiles(
    urlTemplate = "https://{s}.google.com/vt/lyrs=s&x={x}&y={y}&z={z}",
    options = tileOptions(subdomains = c('mt0', 'mt1', 'mt2', 'mt3')),
    attribution = '&copy; <a href="https://maps.google.com">Google Maps</a>',
    group = "Google Satellite"
  ) %>%
  # 3. Layer Control Toggle
  addLayersControl(
    baseGroups = c("Google Road", "Google Satellite"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  
  # Add the N-S-W-E Compass Legend in the bottom right corner
  addControl(
    html = '<div style="background-color: rgba(44, 62, 80, 0.92); color: #ecf0f1; padding: 8px 12px; border-radius: 6px; font-family: Prompt, sans-serif; text-align: center; box-shadow: 0 2px 8px rgba(0,0,0,0.4); border: 1px solid rgba(255,255,255,0.1); font-size: 11px; line-height: 1.2;">
              <div style="color: #94a3b8; font-size: 9px; font-weight: 600; letter-spacing: 0.5px; margin-bottom: 2px;">COMPASS</div>
              <b style="color: #38bdf8;">N</b><br>
              <span style="letter-spacing: 2px;">W &nbsp; E</span><br>
              <b>S</b>
            </div>',
    position = "bottomright"
  ) %>%
  addScaleBar(
    position = "bottomleft",
    options = scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE)
  ) %>%
  setView(lng = 7.000, lat = 50.887, zoom = 14) %>%
  addCircleMarkers(
    radius = 6,
    color = "#0D0658",
    stroke = TRUE,
    weight = 1,
    fillColor = "#F6271A",
    fillOpacity = 0.8,
    popup = ~paste0(
      "<strong>Incident Details:</strong><br>",
      "-----------------------------------<br>",
      "<strong>Accident Type:</strong> ", traffic_accident, "<br>",
      "<strong>Severity:</strong> ", severity, "<br>",
      "<strong>Year:</strong> ", year, "<br>",
      "<strong>Month:</strong> ", Months, "<br>",
      "<strong>Time:</strong> ", sprintf("%02d:00", hour), "<br>",
      "<strong>Lighting:</strong> ", light_cond, "<br>",
      "<strong>Road Condition:</strong> ", road_cond
    )
  )

```


Row {data-height=15}
-------------------------------------

### Interactive Bar Chart
```{r}

library(ggplot2)
library(plotly)
library(dplyr)
library(ggplot2)
library(plotly)
library(dplyr)

# 1. Convert the shared data directly to a standard base data frame 
# This completely strips away the sf class and prevents any mutate/format errors
df_clean <- as.data.frame(sd$data(), stringsAsFactors = FALSE)

# 2. Count accidents using your existing Months column safely
monthly_counts <- df_clean %>%
  count(Months, name = "total")

# 3. Build a crisp, high-contrast ggplot
p <- ggplot(monthly_counts, aes(x = Months, y = total, text = paste("Month:", Months, "<br>Accidents:", total))) +
  geom_col(fill = "#38bdf8", width = 0.7) +
  labs(
    title = "Verkehrsunfälle pro Monat (von 2022 bis 2025)",
    x = "Monat",
    y = "Anzahl der Verkehrsunfälle "
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(color = "#2c3e50", face = "bold", size = 14),
    axis.title = element_text(color = "#2c3e50", face = "bold", size = 11),
    axis.text = element_text(color = "#2c3e50", size = 10),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(color = "#e2e8f0", linewidth = 0.5),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA)
  )

# 4. Convert to interactive plotly widget while preserving Crosstalk linkage
ggplotly(p, tooltip = "text") %>%
  config(displayModeBar = FALSE) %>%
  layout(
    paper_bgcolor = "transparent",
    plot_bgcolor = "transparent"
  )

```

































































































































